---
title: "02-analysis"
format: html
---

# Introduction

 In the following notebook I will compare the weather at stadiums on match days.
 
# Libraries

```{r}
# Installing packages
library(tidyverse)
library(lubridate)
```


# Loading data

```{r}
spring_schedule_2025 <- read_csv("data-processed/exact_2025_weather_spring_sched.csv")
summer_schedule_2025 <- read_csv("data-processed/exact_2025_weather_summer_sched.csv")
early_kickoff_schedule_spring <- read_csv("data-processed/weekly_schedule_early_spring.csv")
early_kickoff_schedule_summer <- read_csv("data-processed/weekly_schedule_early_summer.csv")
late_kickoff_schedule_spring <- read_csv("data-processed/weekly_schedule_late_spring.csv")
late_kickoff_schedule_summer <- read_csv("data-processed/weekly_schedule_late_summer.csv")
```


# Count viz

## Making a dataframe with the counts of matchdays ovre 30 degrees

```{r}

spring_over_thirty_count <- early_kickoff_schedule_spring |> 
  mutate(week_year = isoweek(true_date)) |>
  group_by(Team, week_year) |> 
  summarise(avg_temp = mean(feelslike)) |> 
  filter(avg_temp >= 32) |>
  ungroup() |> 
  group_by(Team) |> 
  count() |> 
  ungroup()

summer_over_thirty_count <- early_kickoff_schedule_summer |> 
  mutate(week_year = isoweek(true_date)) |>
  group_by(Team, week_year) |> 
  summarise(avg_temp = mean(feelslike)) |> 
  filter(avg_temp >= 32) |>
  ungroup() |> 
  group_by(Team) |> 
  count() |> 
  ungroup()



spring_under_freezing_count <- early_kickoff_schedule_spring |> 
  mutate(week_year = isoweek(true_date)) |>
  group_by(Team, week_year) |> 
  summarise(avg_temp = mean(feelslike)) |> 
  filter(avg_temp <= 0) |>
  ungroup() |> 
  group_by(Team) |> 
  count() |> 
  ungroup()



summer_under_freezing_count <- early_kickoff_schedule_summer |> 
  mutate(week_year = isoweek(true_date)) |>
  group_by(Team, week_year) |> 
  summarise(avg_temp = mean(feelslike)) |> 
  filter(avg_temp <= 0) |>
  ungroup() |> 
  group_by(Team) |> 
  count() |> 
  ungroup()

missing_teams <- summer_under_freezing_count |>
  filter(!Team %in% spring_under_freezing_count$Team) |>
  mutate(n = 0)

spring_under_freezing_count <- bind_rows(spring_under_freezing_count, missing_teams)

```


## Making the Viz

```{r}
summer_over_thirty_count |> mutate(difference_n = spring_over_thirty_count$n-summer_over_thirty_count$n) |> 
  ggplot(aes(x = difference_n, y = fct_reorder(Team, difference_n))) +
  geom_col() +
  scale_y_discrete(name = NULL) +
  scale_x_continuous(name = NULL)

summer_under_freezing_count |> mutate(difference_n = spring_under_freezing_count$n-summer_under_freezing_count$n) |> 
  ggplot(aes(x = difference_n, y = fct_reorder(Team, difference_n))) +
  geom_col() +
  scale_y_discrete(name = NULL) +
  scale_x_continuous(name = NULL, limits = c(-20,5), breaks = seq(-20,5,5))
```




# Comparing weather across schedule types

```{r}
# Early Kickoff
early_kickoff_comparison <- early_kickoff_schedule_spring |> mutate(temp_difference = early_kickoff_schedule_spring$temp - early_kickoff_schedule_summer$temp, feels_like_difference = early_kickoff_schedule_spring$feelslike - early_kickoff_schedule_summer$feelslike, precip_difference = early_kickoff_schedule_spring$precip - early_kickoff_schedule_summer$precip, snow_difference = early_kickoff_schedule_spring$snow - early_kickoff_schedule_summer$snow)
```

```{r}
# Late Kickoff
late_kickoff_comparison <- late_kickoff_schedule_spring |> mutate(temp_difference = late_kickoff_schedule_spring$temp - late_kickoff_schedule_summer$temp, feels_like_difference = late_kickoff_schedule_spring$feelslike - late_kickoff_schedule_summer$feelslike, precip_difference = late_kickoff_schedule_spring$precip - late_kickoff_schedule_summer$precip, snow_difference = late_kickoff_schedule_spring$snow - late_kickoff_schedule_summer$snow)
```


# Viewing the new data frame

```{r}
early_kickoff_comparison |> sample_n(10)


late_kickoff_comparison |> sample_n(10)
```

# Function to summarize hourly data to daily

```{r}
# Converts empty precipitation values to 0, removes an empty observation
hourly_to_daily <- function(dataframe){
  #Cleaning the date, creating a day column, and converting NA to 0.
  no_na_dataframe <- dataframe |> mutate(date = as.Date(datetime), precipitation = if_else(is.na(dataframe$precip), 0, dataframe$precip),snow = if_else(is.na(dataframe$snow), 0, dataframe$snow), week = isoweek(date))
  #Finding the high temp, low temp and precipitation for each day.
  dataframe_clean <- no_na_dataframe |>  
  group_by(Team, date) |> 
  summarise(temp = mean(temp, na.rm = TRUE),
            feelslike = mean(feelslike, na.rm = TRUE),
            precipitation = sum(precip),
            snow = sum(snow),
            week = week) |> 
  group_by(Team, week) |> 
  summarise(temp = mean(temp, na.rm = TRUE),
            feelslike = mean(feelslike, na.rm = TRUE),
            precipitation = sum(precipitation),
            snow = sum(snow))  

  
  return(dataframe_clean)
}
```

```{r}
summarized_spring_sched_2025 <- hourly_to_daily(spring_schedule_2025)
summarized_summer_sched_2025 <- hourly_to_daily(summer_schedule_2025)
```



# Summarizing Data frames

```{r}
early_kickoff_comp_summary <- hourly_to_daily(early_kickoff_comparison) |> mutate(matchweek = isoweek(date) - 7)

late_kickoff_comp_summary <- hourly_to_daily(late_kickoff_comparison) |> mutate(matchweek = isoweek(date) - 7)
```

# Analyzing dataframes

```{r}
early_kickoff_comp_summary |> head()

late_kickoff_comp_summary |> head()
```

```{r}
early_kickoff_comp_summary |> 
  group_by(Team,matchweek) |> 
  summarise(avg_temp_diff = mean(temp_diff),
            avg_feels_like_diff = mean(feels_like_diff)) |> 
  filter(matchweek == 1) |> arrange(desc(avg_feels_like_diff))

```

# Comparing change in hot days 